PANCHANGE_descriptives

### 2020-11-06 PANCHANGE_exclusion_descriptives.Rmd

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(polycor)
library(corrplot)
library(patchwork)
library(broom)
library(tidyverse)
library(ggpubr)
library(stringr)
```

@@

## Read in data
```{r read in data}
dat.raw <- readRDS(file = paste0(data_path, "four_cohorts_variables.rds"))
dim(dat.raw)
colnames(dat.raw)
```

@@

Exclude GLAD/EDGI individuals that didn't take part in COPING
Exclude individuals with NA for the MHD categories
Individuals with NAs for age/gender have already been excluded in pre-processing script, but they get excluded here to be safe and to double check previous programming.

@@

# Data availability per questionnaire GAD, PHQ, PCL, and OCI-R

Data availability GAD
```{r gad availability}
dat.raw %>%
  freq(
    data_group_gad,
    cumul = F)
```

Data availability PHQ
```{r phq availability}
dat.raw %>%
  freq(
    data_group_phq,
    cumul = F)
```

Data availability PCL
```{r pcl availability}
dat.raw %>%
  freq(
    data_group_pcl,
    cumul = F)
```

Data availability OCI-R
```{r ocir availability}
dat.raw %>%
  freq(
    data_group_ocir,
    cumul = F)
```

@@

#Impute no change prepandemic scores

```{r Impute for gad measure}

#Visualise the gad summary scores for baseline and retro
head(dat$gad.sum_score_base, 30)
head(dat$gad.sum_score_retro, 30)

#Visualise the gad pandemic feelings variables
head(dat$gad.pandemic_felt_feelings, 10)
head(dat$gad.pandemic_felt_feelings_numeric, 10)

#Edit the gad sum retro score as follows: if the pt reports no different since pandemic and there is a missing value for their retro score, impute their baseline score, else if there isn't a missing value for their retro score, leave it as it is.

impute_dat <- dat %>% 
  mutate(gad.sum_score_retro =
           case_when(gad.pandemic_felt_feelings == "No different" & is.na(gad.sum_score_retro)                      ~ gad.sum_score_base,
                     !is.na(gad.sum_score_retro) ~ gad.sum_score_retro))

#Check that the Na's have been replaced with duplicated baseline sum scores
head(impute_dat$gad.sum_score_base, 30)
head(impute_dat$gad.sum_score_retro, 30)

```{r Impute for phq measure}

#Visualise the phq summary scores for baseline and retro
head(dat$phq.sum_score_base, 30)
head(dat$phq.sum_score_retro, 30)

#Visualise the phq pandemic feelings variables
head(dat$phq.pandemic_felt_feelings, 10)
head(dat$phq.pandemic_felt_feelings_numeric, 10)

#Edit the phq sum retro score as follows: if the pt reports no different since pandemic and there is a missing value for their retro score, impute their baseline score, else if there isn't a missing value for their retro score, leave it as it is.

impute_dat <- dat %>% 
  mutate(phq.sum_score_retro =
           case_when(phq.pandemic_felt_feelings == "No different" & is.na(phq.sum_score_retro)                      ~ phq.sum_score_base,
                     !is.na(phq.sum_score_retro) ~ phq.sum_score_retro))

#Check that the Na's have been replaced with duplicated baseline sum scores
head(impute_dat$phq.sum_score_base, 30)
head(impute_dat$phq.sum_score_retro, 30)

#Baseline survey completion (N per month)

```{r N completing coping per month}

#Visualise some dates when participants completed the questionnaire
head(dat$startDate.coping, 20)

#Divide the data by sign up per month and summarise the amount of pt's completing the survey that month

april_sign_up <- dat[str_detect(dat$startDate.coping, "-04-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(April = n())

may_sign_up <- dat[str_detect(dat$startDate.coping, "-05-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(May = n())

june_sign_up <- dat[str_detect(dat$startDate.coping, "-06-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(June = n())

july_sign_up <- dat[str_detect(dat$startDate.coping, "-07-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(July = n())

august_sign_up <- dat[str_detect(dat$startDate.coping, "-08-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(August = n())

september_sign_up <- dat[str_detect(dat$startDate.coping, "-09-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(September = n())

#Merge results into one tibble

coping_sign_up <- april_sign_up %>% 
  cbind(may_sign_up, june_sign_up, july_sign_up, august_sign_up, september_sign_up)

#Gather the variables
coping_sign_up <- coping_sign_up %>% 
  gather(key = "month", value = "sign_up_number")

#Proportion calculation
coping_sign_up <- coping_sign_up %>% 
  mutate(percentage = 
           round((sign_up_number / sum(sign_up_number) * 100), 2))

coping_sign_up

@@

#Correlation between retropsective impute and prepan

```{r Correlation prep chunk}
# Select items for correlation matrix
prepan_retro_items <- c(
    "gad.sum_score_prepan",
    "gad.sum_score_retro", 
    "phq.sum_score_prepan",
    "phq.sum_score_retro"
)

#Create a data set with only the COPING samples
prepan_retro_dat <- dat[str_detect(dat$Sample, "NBR|GLAD|EDGI"), ]

#Check that RAMP has been removed
prepan_retro_dat %>% 
  freq(Sample)

# Calculate correlation matrix on specific columns (variables) from your data frame
prepan_retro.corr.mat <- hetcor(
  as.data.frame(prepan_retro_dat[,prepan_retro_items]), # Specify the data
  use = "pairwise.complete.obs" # Use all pairwise observations
)

# Print the correlation matrix
print(prepan_retro.corr.mat, digits = max(3, getOption("digits") - 3))

# Save the correlations from the correlation matrix in an object
prepan_retro.corr.matrix <- as.matrix(prepan_retro.corr.mat)

# Save the p values from the correlation matrix in an object
prepan_retro.p.matrix <- as.matrix(prepan_retro.corr.mat$tests)
```

```{r Correlation matrix of prepan base sum scores}
corrplot(prepan_retro.corr.matrix, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorRampPalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = prepan_retro.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )
```

@@

@@

# highest_education
## Proportion tables
```{r highest_education per Sample prop per total}
highest_education_per_Sample_prop_total <- dat %>%
  count(highest_education, Sample) %>%
  drop_na() %>% 
  mutate(Prop = round(n/sum(n), 2))

@@

```{r highest_education per Sample prop per Sample}
highest_education_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  drop_na() %>% 
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r highest_education per Sample prop cross table}
highest_education_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  drop_na() %>% 
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

```{r highest_education per Sample count cross table}
highest_education_per_Sample_count_cross_table <- dat %>%
  count(highest_education, Sample) %>%
  drop_na() %>% 
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )

>>>>

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(polycor)
library(corrplot)
library(patchwork)
library(broom)
library(tidyverse)
library(ggpubr)
```

@@

## Read in data
```{r read in data}
dat.raw <- readRDS(file = paste0(data_path, "/four_cohorts_variables.rds"))
dim(dat.raw)
colnames(dat.raw)
```

@@

Exclude GLAD/EDGI individuals that didn't take part in COPING
Exclude individuals with NA for the MHD categories
Individuals with NAs for age/gender have already been excluded in pre-processing script, but they get excluded here to be save and to double check previous programming.

@@

# Data availability per questionnaire GAD, PHQ, PCL, and OCI-R
++MD: These 'data availability' chunks caused my R to crash, so I'm hashing them out for now. Others can test if it works for them and unhash as needed, otherwise I think they should be removed.

Data availability GAD
```{r gad availability}
#dat.raw %>%
#  freq(
#    data_group_gad,
#    cumul = F)
```

Data availability PHQ
```{r phq availability}
#dat.raw %>%
#  freq(
#    data_group_phq,
#    cumul = F)
```

Data availability PCL
```{r pcl availability}
#dat.raw %>%
#  freq(
#    data_group_pcl,
#    cumul = F)
```

Data availability OCI-R
```{r ocir availability}
#dat.raw %>%
#  freq(
#    data_group_ocir,
#    cumul = F)
```

@@

@@

@@

```{r Disorder_inc by Sample prop table}
Disorder_inc_by_Sample_collapsed <- dat.long.psych %>%
  group_by(Disorder_inc) %>%
  filter(Diagnosis == 1) %>%
  count(Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(Disorder_inc_by_Sample_collapsed)
```

@@

# highest_education
## Proportion tables
```{r highest_education per Sample prop per total}
highest_education_per_Sample_prop_total <- dat %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r highest_education per Sample prop per Sample}
highest_education_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r highest_education per Sample prop cross table}
highest_education_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

```{r highest_education per Sample count cross table}
highest_education_per_Sample_count_cross_table <- dat %>%
  count(highest_education, Sample) %>%
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )
